Learning \strips\ Action Models from State-Constraints

### 2017-12-07 doc/ijcai-18/constraints-learning.tex
building first version ijcai18 paper

\begin{abstract}
This paper presents a classical planning compilation for learning \strips\ action models from observations of plan executions. Remarkably the compilation is fed with {\em state-constraints}, that bound the space of possible action models, and does not require the precise actions in the plan executions. A plan that solves the classical planning task resulting from this compilation induces a \strips\ action model that is compliant with the given state-constraints.  When the amount of input knowledge is too scarce, e.g. the state-constraints are too loose, evaluating learned \strips\ models with respect to reference action model becomes hard. The paper introduces a novel evaluation methodology that allows to asses the quality of learned \strips\ models even if actions are reformulated. Last but not least, the paper shows the flexibility of our compilation to include further information about the observed plan executions and increase the accuracy of the learned models.   
\end{abstract}

@@

Learning action models from observations of plan executions is a promising alternative to the hand-coding of planning action models. This is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}. Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models from observations of plan executions that:
\begin{enumerate}
\item Can be defined as a classical planning compilation which opens the door to the bootstrapping of planning action models.
\item Does not require the precise actions applied in the observed plan executions but is extensible to further information about the observed plan executions to improve the learned models.
\item Assesses the quality of learned \strips\ models with a reference model even when the learning hypothesis space is so low constrained that the learned actions models can be reformulated and still be compliant with the inputs. %For instance in the {\em blocksworld}, given a low amount of input knowledge, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa) making non trivial evaluating the learned model with respect to a reference model.
\end{enumerate}

@@

\begin{figure}
\begin{footnotesize}
\begin{verbatim}
(:action stack
  :parameters (?v1 ?v2 - object)
  :precondition (and (holding ?v1) (clear ?v2))
  :effect (and (not (holding ?v1))
               (not (clear ?v2))
               (handempty) (clear ?v1)
               (on ?v1 ?v2)))
\end{verbatim}
\end{footnotesize}
 \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
\label{fig:stack}
\end{figure}

@@

\subsection{State-constraints}
A {\em state invariant} is a formula $\phi$ that holds in the initial state of a given classical planning problem, $I\models \phi$, and at any state reachable from that state, i.e. for all states $s$ reachable from $I$ it holds that $s\models \phi$. State invariants are traditionally useful for making satisfiability planning and backward search more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.

A {\em mutex} (meaning mutually exclusive) is a particular case of state invariant that takes the form of a binary clause and that represents a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}. For instance in a three-blocks blocksworld, $on(block1,block2)$ and $on(block1,block3)$ are mutex because $block1$ can only be on top of a single block.

A {\em lifted invariant} (also called schematic invariants in the literature) is a state invariant defined using a first order formula over the predicates of a given planning problem~\cite{rintanen:schematicInvariants:AAAI2017}. A {\em domain invariant} is a state invariant that holds for any possible initial state, they are instance-independent and often, they are lifted invariants~\cite{fox:TIM:JAIR1998}.  For instance in the blocksworld, $\forall x,y,z:\ (on(x,y)\wedge on(x,z)) \implies y=z$, is a domain and lifted invariant because a block can only be on top of a single block no matter the initial state of the classical planning problem and no matter the particular identity of the blocks.

@@

We formalize a more challenging learning task, where less input knowledge is available. This learning task, denoted by $\Lambda=\tup{\Psi,\Sigma,\Phi}$, corresponds to observing an agent acting in the world but watching only the results of its plan executions:
\begin{itemize}
\item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}. Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
\item $\Phi$ is a set of {\em lifted domain mutex}.
\end{itemize}

A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, the given states sequences in $\Sigma$ and the set of state constraints $\Phi$. In this learning scenario, a solution must not only determine a possible \strips\ action model but also the content of the {\em unobserved} plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain $\Sigma$ and $\Phi$ with the learned \strips\ model $\Xi$.

@@

\vspace{0.2cm}

\begin{subfigure}{.6\textwidth}
{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
\begin{lstlisting}[mathescape]
\end{lstlisting}
\vspace{0.1cm}
\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
\node (initD) [draw]{D};
\node (initC) [draw, above=of initD.north]{C};
\node (initB) [draw, above=of initC.north]{B};
\node (initA) [draw, above=of initB.north]{A};
\draw[thick] (-1,-0.25) -- (2.5,-0.25);

\node (goalA) [draw, right=10mm of initD]{A};
\node (goalB) [draw, right=10mm of initC]{B};
\node (goalC) [draw, right=10mm of initB]{C};
\node (goalD) [draw, right=10mm of initA]{D};
\end{tikzpicture}
\vspace{0.6cm}
\end{subfigure}%

\vspace{0.2cm}

\begin{footnotesize}
\begin{subfigure}{.6\textwidth}
{\tt ;;; Lifted domain mutexes in $\Phi$}
\begin{verbatim}
(:derived (invariant-1-1)
  (forall (?o1 - object)
      (not (and (handempty) (holding ?o1)))))

(:derived (invariant-1-2)
  (forall (?o1 - object)
      (not (and (holding ?o1) (clear ?o1)))))

(:derived (invariant-1-3)
  (forall (?o1 - object)
      (not (and (holding ?o1) (ontable ?o1)))))

(:derived (invariant-1-4)
  (forall (?o1 - object)
      (not (and (on ?o1 ?o1)))))

(:derived (invariant-2-1)
  (forall (?o1 ?o2 - object)
      (not (and (on ?o1 ?o2) (holding ?o1)))))

(:derived (invariant-2-2)
  (forall (?o1 ?o2 - object)
      (not (and (on ?o1 ?o2) (holding ?o2)))))

(:derived (invariant-2-3)
  (forall (?o1 ?o2 - object)
      (not (and (on ?o1 ?o2) (clear ?o2)))))

(:derived (invariant-2-4)
  (forall (?o1 ?o2 - object)
      (not (and (on ?o1 ?o2) (ontable ?o1)))))

(:derived (invariant-2-5)
  (forall (?o1 ?o2 - object)
      (not (and (on ?o1 ?o2) (on ?o2 ?o1)))))
\end{verbatim}
\end{subfigure}%
\end{footnotesize}

 \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
\label{fig:lexample}
\end{figure}

\subsection{Compiling the learning of \strips\ action models into classical planning}

Our approach for addressing the learning task $\Lambda$, is compiling it into a classical planning task with conditional effects. The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item Programs the \strips\ action model $\Xi$. A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (states $\Sigma$ and formulae $\Phi$). For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying every $\phi\in\Phi$ at every reached state. We call this process the validation of the programmed \strips\ action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}. %If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
\end{enumerate}

@@

Now we are ready to formalize the compilation. Given a learning task $\Lambda=\tup{\Psi,\Sigma,\Phi}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
\begin{itemize}
\item $F_{\Lambda}$ extends $F$ with:
\begin{itemize}
\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$. If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$. For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$}.
\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is being validated.
\end{itemize}
\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label) and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true. Our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect.
\item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
\item $A_{\Lambda}$ contains actions of three kinds:
\begin{enumerate}
\item Actions for {\em programming} an operator schema $\xi\in\Xi$:
\begin{itemize}
\item Actions for {\bf removing} a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.

@@

\section{Constraining the learning hypothesis space with additional input knowledge}
\label{sec:Constraining}
Here we show that further input knowledge can be used to constrain the space of possible action models and make the learning of \strips\ action models more practicable.

\subsection{State constraints}
Every state invariant $\phi\in\Phi$ is added as an extra precondition of actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema. Likewise every state invariant $\phi\in\Phi$ must hold at the last reached state so they are included aas extra goals in $G_{\Lambda}$.

With regard to state trajectories. Now $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is not a set of $(initial, final)$ state pairs but a set of state trajectories $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, that comprises the sequence of states resulting from executing the {\em unobserved} plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.

\begin{itemize}
\item Fluents, $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are added to $F_{\Lambda}$ with the aim of iterating through state trajectories. Likewise $I_{\Lambda}$ includes also now the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$} 
\item Actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema have an extra conditional effect $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$. Likewise actions $\mathsf{validate_{t}}$ for {\em validating} a learning example {\tt\small $1\leq t\leq \tau$} add $at_1$ and delete $at_{|\pi_t|}$.
\end{itemize}

@@

Here we augment the input knowledge with the actions executed by the observed agent, when available. So the leaerning task is now defined as $\Lambda=\tup{\Psi,\Sigma,\Phi,\Pi}$

@@

We extend the compilation to consider the actions in the executed plans. Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Phi,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
\begin{itemize}
\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$. 
\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$. Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
\item With respect to $A_{\Lambda'}$.
\begin{enumerate}
\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step. This mechanism forces that these actions are only applied as in the example plans.
\item The actions for {\em validating} the current learning example have an extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
\begin{small}
\begin{align*}
&\{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
\end{align*}
\end{small}
\end{enumerate}
\end{itemize}

@@

\subsection{Learning with state-constraints}
For each domain we provide a set of {\em lifted domain mutex},  lifted domain invariants that takes the form of a binary clauses. For instance, in the blocksworld $\forall x,y$ then $clear(x)$ and $on(x,y)$ is a {\em lifted domain mutex}.

@@

\section{Conclusions}
{\em Linear Temporal Logic} (LTL) also includes model operator that also allows to represent state constraints~\cite{haslum:LTL:ecai10}. For instance the {\em always} operator, denoted by $\square$, defines constraints that, like {\em lifted domain invariants}, must be true at any reachable state.

>>>>

\begin{abstract}
We present a classical planning compilation for learning \strips\ action models from state-constraints. Remarkably the compilation does not require the precise actions in the plan executions, despite they can be included to constraint further the space of possible action models. A plan that solves the classical planning task resulting from our compilation induces a \strips\ action model that is compliant with all the constraints given as input.  The paper also shows that, when input constraints are too loose, evaluating learned \strips\ models with respect to reference model becomes non trivial (actions can be reformulated, i.e. their semantic can be swapped) and introduces a novel evaluation method that asses learned \strips\ models even if they are reformulated.
\end{abstract}

@@

The Machine Learning of planning action models is a promising alternative to hand-coding them with sophisticated algorithms like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}. Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that:
\begin{enumerate}
\item Is defined as a classical planning compilation which opens the door to the bootstrapping of planning action models.
\item Does not require the actions applied in the observed plan executions but accepts this information to improve the quality of the learned models.
\item Assesses learned \strips\ models with respect to a reference model even when the learning hypothesis space is so low constrained that actions can be reformulated and still be compliant with the inputs. %For instance in the {\em blocksworld}, given a low amount of input knowledge, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa) making non trivial evaluating the learned model with respect to a reference model.
\end{enumerate}

@@

\begin{figure}
\begin{scriptsize}
\begin{verbatim}
(:action stack
  :parameters (?v1 ?v2 - object)
  :precondition (and (holding ?v1) (clear ?v2))
  :effect (and (not (holding ?v1)) (not (clear ?v2))
               (handempty) (clear ?v1) (on ?v1 ?v2)))
\end{verbatim}
\end{scriptsize}
 \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
\label{fig:stack}
\end{figure}

@@

\subsection{State-constraints}
A {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ reachable from $I$. State invariants are traditionally useful for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.

A {\em mutex} (mutually exclusive) is a particular state invariant that takes the form of a binary clause and that represents a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}. For instance in a three-blocks blocksworld, $on(block1,block2)$ and $on(block1,block3)$ are mutex because $block1$ can only be on top of a single block.

A {\em lifted invariant} (also called schematic invariants in the literature) is a state invariant defined using a first order formula~\cite{rintanen:schematicInvariants:AAAI2017}. A {\em domain invariant} is a state invariant that is instance-independent (it holds for any possible initial state) and often, they are lifted invariants~\cite{fox:TIM:JAIR1998}.  For instance in the blocksworld, $\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em lifted domain mutex} because the robot hand can never be empty and holding a block at the same time.

@@

We formalize a more challenging learning task, where less input knowledge is available. This learning task, denoted by $\Lambda=\tup{\Psi,\Sigma,\Phi}$, corresponds to observing an agent acting in the world but watching only the results of its plan executions:
\begin{itemize}
\item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}. Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
\item $\Phi$ is a set of {\em lifted domain invariants}.
\end{itemize}

A solution to $\Lambda$ is a set of operator schema $\Xi$ compliant with the predicates in $\Psi$, the labels $\Sigma$, and the set of state-constraints $\Phi$. A planning compilation is a suitable approach for addressing the $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that explain $\Sigma$ and $\Phi$.

@@

\vspace{0.2cm}
\begin{subfigure}{.6\textwidth}
{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
\begin{lstlisting}[mathescape]
\end{lstlisting}
\vspace{-0.2cm}

\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
\node (initD) [draw]{D};
\node (initC) [draw, above=of initD.north]{C};
\node (initB) [draw, above=of initC.north]{B};
\node (initA) [draw, above=of initB.north]{A};
\draw[thick] (-1,-0.25) -- (2.5,-0.25);

\node (goalA) [draw, right=10mm of initD]{A};
\node (goalB) [draw, right=10mm of initC]{B};
\node (goalC) [draw, right=10mm of initB]{C};
\node (goalD) [draw, right=10mm of initA]{D};
\end{tikzpicture}
\end{subfigure}%

\vspace{0.6cm}

\begin{subfigure}{.6\textwidth}
{\tt ;;; Lifted domain invariants in $\Phi$}
\begin{scriptsize}
\begin{verbatim}
 (forall (?o1 - object)
   (not (and (handempty) (holding ?o1)))))

 (forall (?o1 - object)
   (not (and (holding ?o1) (clear ?o1)))))

 (forall (?o1 - object)
   (not (and (holding ?o1) (ontable ?o1))))

 (forall (?o1 - object)
   (not (and (on ?o1 ?o1))))

 (forall (?o1 ?o2 - object)
   (not (and (on ?o1 ?o2) (holding ?o1))))

 (forall (?o1 ?o2 - object)
   (not (and (on ?o1 ?o2) (holding ?o2))))

 (forall (?o1 ?o2 - object)
    (not (and (on ?o1 ?o2) (clear ?o2))))

 (forall (?o1 ?o2 - object)
    (not (and (on ?o1 ?o2) (ontable ?o1))))

 (forall (?o1 ?o2 - object)
    (not (and (on ?o1 ?o2) (on ?o2 ?o1)))))
\end{verbatim}
\end{scriptsize}
\end{subfigure}%
 \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
\label{fig:lexample}
\end{figure}

\subsection{Learning \strips\ action models with classical planning}

Our approach for addressing the learning task $\Lambda$, is compiling it into a classical planning task with conditional effects. The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item Programs the \strips\ action model $\Xi$. A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
\item Validates the programmed \strips\ action model $\Xi$ in the input constraints (labels $\Sigma$ and formulae $\Phi$). For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying every $\phi\in\Phi$ at every reached state. We call this process the validation of the programmed \strips\ action model $\Xi$, at the $t^{th}$ learning example {\small $1\leq t\leq \tau$}. %If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
\end{enumerate}

@@

Now we are ready to formalize the compilation. Given a learning task $\Lambda=\tup{\Psi,\Sigma,\Phi}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
\begin{itemize}
\item $F_{\Lambda}$ extends $F$ with:
\begin{itemize}
\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$. If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$. For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$}.
\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is being validated.
\end{itemize}
\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label) and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true. Our compilation assumes that initially operator schemas are programmed with every possible precondition, no negative effect and no positive effect.
\item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
\item $A_{\Lambda}$ contains actions of three kinds:
\begin{enumerate}
\item Actions for {\em programming} operator schema $\xi\in\Xi$:
\begin{itemize}
\item Actions for {\bf removing} a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.

@@

\section{Constraining the learning hypothesis space}
\label{sec:Constraining}
Here we show that constraints can be used to reduce the space of possible action models and make the learning of \strips\ action models more practicable.

\subsection{State constraints}
Every state invariant $\phi\in\Phi$ is added as an extra precondition of actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema. Likewise, every state invariant $\phi\in\Phi$ is also added as an extra goal in $G_{\Lambda}$ because formulae $\phi\in\Phi$ must hold at every reached state, including the last reached state.

If the state trajectory corresponding to a plan execution is available it can also be included in the compilation. In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state trajectories $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, that comprises the sequence of states resulting from executing the {\em unobserved} plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.

\begin{itemize}
\item Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are added to $F_{\Lambda}$ to iterate through the state trajectories. Likewise $I_{\Lambda}$ includes also now the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$} 
\item Actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema have extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$. Actions $\mathsf{validate_{t}}$ for {\em validating} the $t^{th}$ learning example add $at_1$ and delete $at_{|\pi_t|}$.
\end{itemize}

{\em Linear Temporal Logic} (LTL) allows to represent more expressive state-constraints~\cite{haslum:LTL:ecai10}. For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define state-constraints that, unlike {\em state invariants}, must be true at least at one of the reached states. LTL constraints can be included in our compilation following the ideas for compiling first-order temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) adds as new goals classical planning goals the aceptor states of the equivalent automata.

@@

Here we show that our classical planning compilation is also flexible to include as input constraints, if available, the actions in the plans executed by the observed agent. Now the learning task is defined as $\Lambda=\tup{\Psi,\Sigma,\Phi,\Pi}$, where:

@@

We extend the compilation to consider the actions in the executed plans. Given a learning task $\Lambda=\tup{\Psi,\Sigma,\Phi,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}$ as follows:
\begin{itemize}
\item $F_{\Lambda}$ is extended with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$ and {\small $1\leq j\leq |\pi_t|$}. 
\item $I_{\Lambda}$ is extended with fluents $F_{\pi_1}$. Goals are $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
\item With respect to $A_{\Lambda}$.
\begin{enumerate}
\item The actions for {\em programming} the preconditions/effects of a given schema $\xi\in\Xi$ are the same.
\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step. This mechanism forces that these actions are only applied as in the example plans.
\item The actions for {\em validating} the current learning example have an extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
\begin{small}
\begin{align*}
&\{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
\end{align*}
\end{small}
\end{enumerate}
\end{itemize}

@@

When the learning hypothesis space is low constrained, the learned actions can be reformulated and still be compliant with the inputs. For instance in the {\em blocksworld}, given a low amount of input knowledge, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa) making non trivial to compute {\em precision} and {\em recall} with respect to a reference model. To address this issue we define the following evaluation methodology that deals with action reformulation.

Given a reference \strips\ action model $\Xi^*$ and the learned \strips\ action model $\Xi$ we define these two bijective functions $f_p:\Xi \mapsto \Xi^*$ and $f_r:\Xi \mapsto \Xi^*$ such that $f_p$ and $f_r$ respectively maximize the the accumulated {\em precision} and {\em recall}. With this defined we compute the {\em precision} of an \strips\ action $\xi$ with respect to the action $f_p(\xi)$. Likewise, the {\em recall} of an \strips\ action $\xi$ is computed with respect to the action $f_r(\xi)$.

\subsection{Learning with state-constraints}
For each domain we provide a set of {\em lifted domain invariants}.

@@

\section{Conclusions}

% Commented for blind submission
%\begin{small}
%\subsection*{Acknowledgment}
%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government. Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
%\end{small}
